---
title: "p8105_hw2_HL3859"
author: "Hanrui Li"
date: "2024-09-25"
output: github_document
---

# Problem 1

```{r}
library(tidyverse)

nyc = 
  read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names()

nyc_clean = 
  nyc |>
  select(line, station_name, station_latitude, station_longitude, route1:route11, 
         entry, vending, entrance_type, ada) |>
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

nyc_clean
```

The dataset contains `line` which shows 36 subway lines. `station_name` represents the names of the station. `station_latitude` and `station_longitude` represent the coordinates of the station. `route1` to `route11` show the subway routes served by the station, there are 11 routes. `entry` indicates whether entrance is allowed (TRUE/FALSE). `vending` indicates whether there is a vending machine (YES/NO). `entrance_type` is the type of entrance (e.g. `r unique(nyc_clean$entrance_type)`). `ada` inndicates whether the station is compliant with ADA standards (TRUE/FALSE). I selected relevant columns from the dataset and converted the `entry` variable to a logical variable for easier analysis. After cleaning, the dataset has `r nrow(nyc_clean)` rows and `r ncol(nyc_clean)` columns. This dataset is tidy as each column represents a variable and each row represents a single observation (an entrance/exit).

```{r}
nyc_clean |>
  distinct(line, station_name) |>
  nrow()
```

- There are 465 distinct stations.

```{r}
nyc_clean |>
  filter(ada == "TRUE") |>
  distinct(line, station_name) |>
  nrow()
```

- 84 stations are ADA compliant.

```{r}
nyc_clean |>
  filter(vending == "NO") |>
  summarise(proportion = mean(entry))
```

- 37.70% of station entrances / exits without vending allow entrance.

```{r}
nyc_routes = 
  nyc_clean |>
  mutate(across(starts_with("route"), as.character)) |>
  pivot_longer(
    route1:route11,
    names_to = "route_name", 
    values_to = "route_number") |>
  filter(!is.na(route_number))

nyc_routes

stations_A =
  nyc_routes |>
  filter(route_number == "A") |>
  distinct(line, station_name)

nrow(stations_A)

stations_A |>
  inner_join(nyc_clean |> filter(ada == "TRUE"), by = c("line", "station_name")) |>
  nrow()
```

- There are 60 distinct stations serve the A train.

- Of the stations that serve the A train, 107 are ADA compliant.


# Problem 2

```{r}
library(readxl)
library(dplyr)

trash_wheel = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", skip = 1) |>
  select(-starts_with(".")) |>
  filter(!is.na(Dumpster)) |>
  mutate(`Sports Balls` = as.integer(round(`Sports Balls`)))

trash_wheel
```











